---
title: "Replicating Duflo (2001) with Rfuzzydid"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicating Duflo (2001) with Rfuzzydid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
```

# Introduction

This vignette walks through a concrete replication exercise: re-estimating the
Duflo (2001) INPRES specification with `Rfuzzydid`.

The key idea is to recover the wage return to schooling from cohort-by-district
variation in treatment intensity, using the fuzzy DID estimators in
de Chaisemartin and D'Haultfoeuille (2018).

The dataset is the bundled `inpresdata.dta` file in `inst/extdata`. The file
originates from David Roodman's replication materials
([droodman/Duflo2001](https://github.com/droodman/Duflo2001/blob/main/inpresdata.dta)).
In the companion archive README
([droodman/Duflo-2001](https://github.com/droodman/Duflo-2001)), he notes that
the main 1995 SUPAS file is "not contained in this repository" and links to
that downloadable location.

# Data Preparation

## Loading the Data

```{r load-data}
library(Rfuzzydid)
library(foreign)

# Resolve package root when rendering inside source tree
find_local_pkg_dir <- function() {
  roots <- c(".", "..", "../..")

  for (root in roots) {
    pkg_desc <- file.path(root, "pkg", "DESCRIPTION")
    root_desc <- file.path(root, "DESCRIPTION")

    if (file.exists(pkg_desc)) {
      return(file.path(root, "pkg"))
    }

    if (file.exists(root_desc)) {
      desc <- tryCatch(read.dcf(root_desc), error = function(e) NULL)
      if (!is.null(desc) && "Package" %in% colnames(desc) && identical(desc[1, "Package"], "Rfuzzydid")) {
        return(root)
      }
    }
  }

  NULL
}

# First try installed package location, then source-tree fallback
data_path <- system.file("extdata", "inpresdata.dta", package = "Rfuzzydid")
if (identical(data_path, "")) {
  local_pkg_dir <- find_local_pkg_dir()
  local_path <- if (!is.null(local_pkg_dir)) {
    file.path(local_pkg_dir, "inst", "extdata", "inpresdata.dta")
  } else {
    ""
  }

  if (file.exists(local_path)) {
    data_path <- local_path
  } else {
    stop("Bundled dataset 'inpresdata.dta' was not found in installed package or source-tree extdata.")
  }
}
raw <- read.dta(data_path)
```

## Constructing the Analysis Sample

Following the original design, we keep two cohort windows:

- 1957-1962 (earlier cohort)
- 1968-1972 (later cohort)

The `t` indicator equals 1 for the later cohort.

```{r sample-construction}
# Define cohort groups
cohort0 <- raw$p504thn >= 57 & raw$p504thn <= 62
cohort1 <- raw$p504thn >= 68 & raw$p504thn <= 72

# Keep observations with non-missing outcome and treatment
keep <- !is.na(raw$lhwage) & !is.na(raw$yeduc) & (cohort0 | cohort1)

dat <- raw[keep, c("birthpl", "lhwage", "yeduc", "p504thn")]

# Time indicator: t = 1 for later cohort
dat$t <- as.integer(dat$p504thn >= 68 & dat$p504thn <= 72)
```

## Constructing Group Variables

The original Stata workflow builds "super-groups" based on how schooling
changes within district between the two cohort windows. We follow the same
logic with a chi-square screen and the sign of the change in mean schooling:

```{r group-construction}
pval_threshold <- 0.5
districts <- sort(unique(dat$birthpl))
gstar_map <- setNames(integer(length(districts)), as.character(districts))

for (g in districts) {
  i <- dat$birthpl == g
  sub <- dat[i, ]
  tab <- table(sub$yeduc, sub$t)
  pval <- suppressWarnings(chisq.test(tab)$p.value)
  evol <- mean(sub$yeduc[sub$t == 1]) - mean(sub$yeduc[sub$t == 0])
  if (!is.na(pval) && pval < pval_threshold) {
    gstar_map[[as.character(g)]] <- if (evol > 0) 1L else if (evol < 0) -1L else 0L
  } else {
    gstar_map[[as.character(g)]] <- 0L
  }
}

dat$gstar <- unname(gstar_map[as.character(dat$birthpl)])
```

This produces three groups:

- `gstar = 1`: Districts where education increased (treatment intensity rose)
- `gstar = -1`: Districts where education decreased
- `gstar = 0`: Districts with no significant change

# Two-Arm Estimation

As in the paper's setup, we estimate two arms separately and combine them after:

## Arm 1: Districts with Increasing Education (g* = 1 vs g* = 0)

```{r arm-inc}
arm_inc_data <- dat[dat$gstar >= 0, ]
arm_inc_data$G <- as.integer(arm_inc_data$gstar == 1)

fit_inc <- fuzzydid(
  data = arm_inc_data,
  formula = lhwage ~ yeduc,
  group = "G",
  time = "t",
  did = TRUE,
  tc = TRUE,
  cic = TRUE,
  newcateg = c(5, 8, 11, 14, 1000),
  nose = TRUE,
  backend = "native"
)

knitr::kable(fit_inc$late)
```

## Arm 2: Districts with Decreasing Education (g* = -1 vs g* = 0)

```{r arm-dec}
arm_dec_data <- dat[dat$gstar <= 0, ]
arm_dec_data$G <- as.integer(arm_dec_data$gstar == -1)

fit_dec <- fuzzydid(
  data = arm_dec_data,
  formula = lhwage ~ yeduc,
  group = "G",
  time = "t",
  did = TRUE,
  tc = TRUE,
  cic = TRUE,
  newcateg = c(5, 8, 11, 14, 1000),
  nose = TRUE,
  backend = "native"
)

knitr::kable(fit_dec$late)
```

# Aggregating Results

To aggregate, we weight each arm by:

- its sample share (`p_inc`, `p_dec`)
- its first-stage strength (the DID in schooling, `dD_inc`, `dD_dec`)

The decreasing arm uses `-dD_dec` so both weights are on a comparable
"increase in treatment" scale.

```{r aggregate}
# Treatment intensity weights
dD_inc <- mean(arm_inc_data$yeduc[arm_inc_data$G == 1 & arm_inc_data$t == 1]) -
          mean(arm_inc_data$yeduc[arm_inc_data$G == 1 & arm_inc_data$t == 0]) -
          mean(arm_inc_data$yeduc[arm_inc_data$G == 0 & arm_inc_data$t == 1]) +
          mean(arm_inc_data$yeduc[arm_inc_data$G == 0 & arm_inc_data$t == 0])

dD_dec <- mean(arm_dec_data$yeduc[arm_dec_data$G == 1 & arm_dec_data$t == 1]) -
          mean(arm_dec_data$yeduc[arm_dec_data$G == 1 & arm_dec_data$t == 0]) -
          mean(arm_dec_data$yeduc[arm_dec_data$G == 0 & arm_dec_data$t == 1]) +
          mean(arm_dec_data$yeduc[arm_dec_data$G == 0 & arm_dec_data$t == 0])

p_inc <- mean(dat$gstar == 1)
p_dec <- mean(dat$gstar == -1)

w_inc <- p_inc * dD_inc
w_dec <- p_dec * (-dD_dec)

# Extract point estimates
did_inc <- fit_inc$late$estimate[fit_inc$late$estimator == "W_DID"]
tc_inc  <- fit_inc$late$estimate[fit_inc$late$estimator == "W_TC"]
cic_inc <- fit_inc$late$estimate[fit_inc$late$estimator == "W_CIC"]

did_dec <- fit_dec$late$estimate[fit_dec$late$estimator == "W_DID"]
tc_dec  <- fit_dec$late$estimate[fit_dec$late$estimator == "W_TC"]
cic_dec <- fit_dec$late$estimate[fit_dec$late$estimator == "W_CIC"]

# Weighted aggregation
did_agg <- (did_inc * w_inc + did_dec * w_dec) / (w_inc + w_dec)
tc_agg  <- (tc_inc  * w_inc + tc_dec  * w_dec) / (w_inc + w_dec)
cic_agg <- (cic_inc * w_inc + cic_dec * w_dec) / (w_inc + w_dec)

cat("Aggregate Estimates:\n")
cat(sprintf("DID = %.6f\n", did_agg))
cat(sprintf("TC  = %.6f\n", tc_agg))
cat(sprintf("CIC = %.6f\n", cic_agg))
```

# Comparison with Stata

The following one-shot Stata code starts from `inpresdata.dta` and reproduces
the arm-level and aggregate point estimates:

```stata
clear all
set more off

local pval_threshold 0.5

use inpresdata.dta, clear
keep if lhwage < . & yeduc < . & (inrange(p504thn,57,62) | inrange(p504thn,68,72))

gen byte t = inrange(p504thn,68,72)
gen gstar = .
levelsof birthpl, local(districts)

local pval_threshold = 0.05 
quietly foreach g of local districts {
    tab yeduc t if birthpl == `g', chi2
    scalar p = r(p)
    summarize yeduc if t == 1 & birthpl == `g', meanonly
    scalar m1 = r(mean)
    summarize yeduc if t == 0 & birthpl == `g', meanonly
    scalar m0 = r(mean)
    scalar evol = m1 - m0
    replace gstar = (scalar(evol) > 0 & scalar(p) < `pval_threshold') - (scalar(evol) < 0 & scalar(p) < `pval_threshold') if birthpl == `g'
}

gen byte inc = (gstar == 1)
gen byte dec = (gstar == -1)
summarize inc, meanonly
scalar p_inc = r(mean)
summarize dec, meanonly
scalar p_dec = r(mean)

* Arm g*=1 vs g*=0
preserve
keep if gstar >= 0
gen byte G = (gstar == 1)
fuzzydid lhwage G t yeduc, did tc cic newcateg(5 8 11 14 1000) nose
matrix M = e(b_LATE)
scalar did_inc = M[1,1]
scalar tc_inc  = M[2,1]
scalar cic_inc = M[3,1]

summarize yeduc if G == 1 & t == 1, meanonly
scalar D11 = r(mean)
summarize yeduc if G == 1 & t == 0, meanonly
scalar D10 = r(mean)
summarize yeduc if G == 0 & t == 1, meanonly
scalar D01 = r(mean)
summarize yeduc if G == 0 & t == 0, meanonly
scalar D00 = r(mean)
scalar dD_inc = (D11 - D10) - (D01 - D00)
restore

* Arm g*=-1 vs g*=0
preserve
keep if gstar <= 0
gen byte G = (gstar == -1)
fuzzydid lhwage G t yeduc, did tc cic newcateg(5 8 11 14 1000) nose
matrix M = e(b_LATE)
scalar did_dec = M[1,1]
scalar tc_dec  = M[2,1]
scalar cic_dec = M[3,1]

summarize yeduc if G == 1 & t == 1, meanonly
scalar D11m = r(mean)
summarize yeduc if G == 1 & t == 0, meanonly
scalar D10m = r(mean)
summarize yeduc if G == 0 & t == 1, meanonly
scalar D01m = r(mean)
summarize yeduc if G == 0 & t == 0, meanonly
scalar D00m = r(mean)
scalar dD_dec_pos = -((D11m - D10m) - (D01m - D00m))
restore

scalar w_inc = p_inc * dD_inc
scalar w_dec = p_dec * dD_dec_pos

scalar did_agg = (did_inc * w_inc + did_dec * w_dec) / (w_inc + w_dec)
scalar tc_agg  = (tc_inc  * w_inc + tc_dec  * w_dec) / (w_inc + w_dec)
scalar cic_agg = (cic_inc * w_inc + cic_dec * w_dec) / (w_inc + w_dec)

display "Arm g*=1: DID = " %9.6f did_inc " ; TC = " %9.6f tc_inc " ; CIC = " %9.6f cic_inc
display "Arm g*=-1: DID = " %9.6f did_dec " ; TC = " %9.6f tc_dec " ; CIC = " %9.6f cic_dec
display "AGG DID = " %9.6f did_agg
display "AGG TC  = " %9.6f tc_agg
display "AGG CIC = " %9.6f cic_agg
```

Running that code in Stata reports:

- Arm `g*=1`: DID = `0.123679191`, TC = `0.079868219`, CIC = `0.074316049`
- Arm `g*=-1`: DID = `0.117024113`, TC = `0.109567707`, CIC = `0.114301855`
- Aggregate: DID = `0.122244473`, TC = `0.086270906`, CIC = `0.082936285`

The next chunk checks parity directly in R:

```{r stata-parity-check}
stata_vals <- c(
  did_inc = 0.123679191,
  tc_inc = 0.079868219,
  cic_inc = 0.074316049,
  did_dec = 0.117024113,
  tc_dec = 0.109567707,
  cic_dec = 0.114301855,
  did_agg = 0.122244473,
  tc_agg = 0.086270906,
  cic_agg = 0.082936285
)

r_vals <- c(
  did_inc = did_inc,
  tc_inc = tc_inc,
  cic_inc = cic_inc,
  did_dec = did_dec,
  tc_dec = tc_dec,
  cic_dec = cic_dec,
  did_agg = did_agg,
  tc_agg = tc_agg,
  cic_agg = cic_agg
)

cmp <- data.frame(
  estimate = names(stata_vals),
  R = unname(r_vals),
  Stata = unname(stata_vals),
  abs_diff = abs(unname(r_vals) - unname(stata_vals))
)

knitr::kable(cmp, digits = 6)

```

Point estimates match to numerical tolerance. Confidence intervals can still
differ when bootstrap is enabled, because draws and RNG streams differ across
software.

# References

Duflo, E. (2001). Schooling and Labor Market Consequences of School Construction in Indonesia. *American Economic Review*, 91(4): 795-813.

de Chaisemartin, C. and D'Haultfoeuille, X. (2018). [Fuzzy Differences-in-Differences](https://doi.org/10.1093/restud/rdx049). *Review of Economic Studies*, 85(2): 999-1028.

de Chaisemartin, C., D'Haultfoeuille, X., and Guyonvarch, Y. (2018). [Fuzzy Differences-in-Differences with Stata](https://doi.org/10.1177/1536867X19854019). *Stata Journal*.
