---
title: "Replicating Duflo (2001) with Rfuzzydid"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicating Duflo (2001) with Rfuzzydid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
```

# Introduction

This vignette demonstrates how to use **Rfuzzydid** to replicate the Duflo (2001) school construction program analysis using fuzzy difference-in-differences, following the methodology of de Chaisemartin and D'Haultfoeuille (2018).

The data comes from the INPRES school construction program in Indonesia, examining the effect of education on wages. We use the bundled `inpresdata.dta` file shipped in `inst/extdata`, sourced from [droodman/Duflo2001](https://github.com/droodman/Duflo2001/blob/main/inpresdata.dta).

# Data Preparation

## Loading the Data

```{r load-data}
library(Rfuzzydid)
library(foreign)

# Load bundled INPRES data from the installed package
data_path <- system.file("extdata", "inpresdata.dta", package = "Rfuzzydid")
if (identical(data_path, "")) {
  stop("Bundled dataset 'inpresdata.dta' was not found in extdata.")
}
raw <- read.dta(data_path)
```

## Constructing the Analysis Sample

Following the original analysis, we restrict to cohorts 1957-1962 (control) and 1968-1972 (treated):

```{r sample-construction}
# Define cohort groups
cohort0 <- raw$p504thn >= 57 & raw$p504thn <= 62
cohort1 <- raw$p504thn >= 68 & raw$p504thn <= 72

# Keep observations with non-missing outcome and treatment
keep <- !is.na(raw$lhwage) & !is.na(raw$yeduc) & (cohort0 | cohort1)

dat <- raw[keep, c("birthpl", "lhwage", "yeduc", "p504thn")]

# Time indicator: t = 1 for later cohort
dat$t <- as.integer(dat$p504thn >= 68 & dat$p504thn <= 72)
```

## Constructing Group Variables

The fuzzy DID methodology constructs super-groups based on whether education evolved differently across districts. We use chi-square tests to classify districts:

```{r group-construction}
pval_threshold <- 0.5
districts <- sort(unique(dat$birthpl))
gstar_map <- setNames(integer(length(districts)), as.character(districts))

for (g in districts) {
  i <- dat$birthpl == g
  sub <- dat[i, ]
  tab <- table(sub$yeduc, sub$t)
  pval <- suppressWarnings(chisq.test(tab)$p.value)
  evol <- mean(sub$yeduc[sub$t == 1]) - mean(sub$yeduc[sub$t == 0])
  if (!is.na(pval) && pval < pval_threshold) {
    gstar_map[[as.character(g)]] <- if (evol > 0) 1L else if (evol < 0) -1L else 0L
  } else {
    gstar_map[[as.character(g)]] <- 0L
  }
}

dat$gstar <- unname(gstar_map[as.character(dat$birthpl)])
```

This creates three super-groups:

- `gstar = 1`: Districts where education increased (treatment intensity rose)
- `gstar = -1`: Districts where education decreased
- `gstar = 0`: Districts with no significant change

# Two-Arm Estimation

The analysis estimates effects separately for two treatment arms, then aggregates:

## Arm 1: Districts with Increasing Education (g* = 1 vs g* = 0)

```{r arm-inc}
arm_inc_data <- dat[dat$gstar >= 0, ]
arm_inc_data$G <- as.integer(arm_inc_data$gstar == 1)

fit_inc <- fuzzydid(
  data = arm_inc_data,
  formula = lhwage ~ yeduc,
  group = "G",
  time = "t",
  did = TRUE,
  tc = TRUE,
  cic = TRUE,
  newcateg = c(5, 8, 11, 14, 1000),
  nose = TRUE,
  backend = "native"
)

knitr::kable(fit_inc$late)
```

## Arm 2: Districts with Decreasing Education (g* = -1 vs g* = 0)

```{r arm-dec}
arm_dec_data <- dat[dat$gstar <= 0, ]
arm_dec_data$G <- as.integer(arm_dec_data$gstar == -1)

fit_dec <- fuzzydid(
  data = arm_dec_data,
  formula = lhwage ~ yeduc,
  group = "G",
  time = "t",
  did = TRUE,
  tc = TRUE,
  cic = TRUE,
  newcateg = c(5, 8, 11, 14, 1000),
  nose = TRUE,
  backend = "native"
)

knitr::kable(fit_dec$late)
```

# Aggregating Results

Following the paper, results are aggregated using treatment-intensity weights:

```{r aggregate}
# Treatment intensity weights
dD_inc <- mean(arm_inc_data$yeduc[arm_inc_data$G == 1 & arm_inc_data$t == 1]) -
          mean(arm_inc_data$yeduc[arm_inc_data$G == 1 & arm_inc_data$t == 0]) -
          mean(arm_inc_data$yeduc[arm_inc_data$G == 0 & arm_inc_data$t == 1]) +
          mean(arm_inc_data$yeduc[arm_inc_data$G == 0 & arm_inc_data$t == 0])

dD_dec <- mean(arm_dec_data$yeduc[arm_dec_data$G == 1 & arm_dec_data$t == 1]) -
          mean(arm_dec_data$yeduc[arm_dec_data$G == 1 & arm_dec_data$t == 0]) -
          mean(arm_dec_data$yeduc[arm_dec_data$G == 0 & arm_dec_data$t == 1]) +
          mean(arm_dec_data$yeduc[arm_dec_data$G == 0 & arm_dec_data$t == 0])

p_inc <- mean(dat$gstar == 1)
p_dec <- mean(dat$gstar == -1)

w_inc <- p_inc * dD_inc
w_dec <- p_dec * (-dD_dec)

# Extract point estimates
did_inc <- fit_inc$late$estimate[fit_inc$late$estimator == "W_DID"]
tc_inc  <- fit_inc$late$estimate[fit_inc$late$estimator == "W_TC"]
cic_inc <- fit_inc$late$estimate[fit_inc$late$estimator == "W_CIC"]

did_dec <- fit_dec$late$estimate[fit_dec$late$estimator == "W_DID"]
tc_dec  <- fit_dec$late$estimate[fit_dec$late$estimator == "W_TC"]
cic_dec <- fit_dec$late$estimate[fit_dec$late$estimator == "W_CIC"]

# Weighted aggregation
did_agg <- (did_inc * w_inc + did_dec * w_dec) / (w_inc + w_dec)
tc_agg  <- (tc_inc  * w_inc + tc_dec  * w_dec) / (w_inc + w_dec)
cic_agg <- (cic_inc * w_inc + cic_dec * w_dec) / (w_inc + w_dec)

cat("Aggregate Estimates:\n")
cat(sprintf("DID = %.6f\n", did_agg))
cat(sprintf("TC  = %.6f\n", tc_agg))
cat(sprintf("CIC = %.6f\n", cic_agg))
```

# Comparison with Stata

The equivalent Stata code is available in the [Duflo2001 repository](https://github.com/droodman/Duflo2001). **Note:** Point estimates from R and Stata will be identical, but bootstrap confidence intervals will differ due to RNG differences between platforms. Results remain comparable across implementations.

# References

Duflo, E. (2001). Schooling and Labor Market Consequences of School Construction in Indonesia. *American Economic Review*, 91(4): 795-813.

de Chaisemartin, C. and D'Haultfoeuille, X. (2018). [Fuzzy Differences-in-Differences](https://academic.oup.com/restud/article/85/2/999/4096388). *Review of Economic Studies*, 85(2): 999-1028.

de Chaisemartin, C., D'Haultfoeuille, X., and Guyonvarch, Y. (2018). [Fuzzy Differences-in-Differences with Stata](https://sites.google.com/site/clementdechaisemartin/statapaper_fuzzydid.pdf). *Stata Journal*.
