---
title: "Stata Parity: Replicating Stata fuzzydid in R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Stata Parity: Replicating Stata fuzzydid in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r parity-helpers, include = FALSE}
find_parity_dir <- function() {
  candidates <- c("tests/testthat", "pkg/tests/testthat", "../tests/testthat")
  for (path in candidates) {
    if (dir.exists(path)) return(path)
  }
  stop("Could not locate parity golden directory.", call. = FALSE)
}

parity_golden_dir <- find_parity_dir()

read_parity_golden <- function(filename) {
  read.csv(file.path(parity_golden_dir, filename), check.names = FALSE)
}

# Prefer local package sources when rendering inside the repository so
# parity checks target the code under development.
find_local_pkg_dir <- function() {
  roots <- c(".", "..", "../..")

  for (root in roots) {
    pkg_desc <- file.path(root, "pkg", "DESCRIPTION")
    root_desc <- file.path(root, "DESCRIPTION")

    if (file.exists(pkg_desc)) {
      return(file.path(root, "pkg"))
    }

    if (file.exists(root_desc)) {
      desc <- tryCatch(read.dcf(root_desc), error = function(e) NULL)
      if (!is.null(desc) && "Package" %in% colnames(desc) && identical(desc[1, "Package"], "Rfuzzydid")) {
        return(root)
      }
    }
  }

  NULL
}

local_pkg_dir <- find_local_pkg_dir()

if (!is.null(local_pkg_dir) && requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(local_pkg_dir, helpers = FALSE, export_all = FALSE, quiet = TRUE)
} else {
  library(Rfuzzydid)
}
```

# Introduction

This vignette demonstrates how **Rfuzzydid** replicates the behavior of the Stata `fuzzydid` command. We use simulated data to show that the R implementation produces identical point estimates and properly handles the same design restrictions as the original Stata package.

The methodology implemented in both packages comes from de Chaisemartin and D'Haultfoeuille (2018), who propose estimators for local average and quantile treatment effects in fuzzy difference-in-differences designs.


# Simulated Data Example

We generate a simple two-period, two-group fuzzy DID design. The data generating process creates:

- A control group (G = 0) observed in periods T = 0 and T = 1
- A treatment group (G = 1) observed in periods T = 0 and T = 1
- Treatment uptake D that increases more in the treatment group between periods
- Outcome Y that responds to treatment

```{r sim-data}
set.seed(50321)

n_cell <- 120  # observations per cell

# Helper function to create each (G, T) cell
make_cell <- function(g, t, prob_d) {
  d <- rbinom(n_cell, size = 1, prob = prob_d)
  # Outcome: base + group effect + time effect + treatment effect + noise
  y <- 1 + 0.5 * g + 0.4 * t + 2.0 * d + sin(seq_len(n_cell) / 7)
  data.frame(y = y, g = g, t = t, d = d)
}

# Build the four cells
df <- rbind(
  make_cell(g = 0, t = 0, prob_d = 0.20),  # Control group, baseline
  make_cell(g = 0, t = 1, prob_d = 0.40),  # Control group, follow-up  
  make_cell(g = 1, t = 0, prob_d = 0.30),  # Treatment group, baseline
  make_cell(g = 1, t = 1, prob_d = 0.80)   # Treatment group, follow-up
)

# Verify cell sizes and treatment rates
aggregate(cbind(n = d) ~ g + t, data = df, FUN = function(x) c(n = length(x), mean = mean(x)))

# Save fixture used by most examples for direct Stata comparison
if (requireNamespace("haven", quietly = TRUE)) {
  haven::write_dta(df, "stata-parity-sim-fixture.dta")
}
```

The treatment rate increases from 30% to 80% in the treatment group, while only increasing from 20% to 40% in the control group. This differential increase creates the "fuzzy" setting.

# Basic Estimation

## Wald-DID, Wald-TC, and Wald-CIC

We request all three main estimators:

```{r basic-est}
fit <- fuzzydid(
  data = df,
  formula = y ~ d,
  group = "g",
  time = "t",
  did = TRUE,
  tc = TRUE,
  cic = TRUE,
  breps = 50
)

summary(fit)

stata_basic <- c(
  W_DID = 2.00000000000000,
  W_TC = 1.97320415000280,
  W_CIC = 2.22209836585117
)
r_basic <- setNames(fit$late$estimate, fit$late$estimator)[names(stata_basic)]
basic_cmp <- data.frame(
  estimator = names(stata_basic),
  R = unname(r_basic),
  Stata = unname(stata_basic),
  abs_diff = abs(unname(r_basic) - unname(stata_basic))
)
knitr::kable(basic_cmp, digits = 8)
```

The output shows point estimates, bootstrap standard errors, and 95% percentile bootstrap confidence intervals for each estimator.

Equivalent Stata code:

```stata
fuzzydid y g t d, did tc cic breps(50)
matrix list e(b_LATE)
matrix list e(se_LATE)
matrix list e(ci_LATE)
```

## Extracting Results Programmatically

The `tidy()` method returns a data frame suitable for downstream analysis:

```{r tidy-output}
library(modelsummary)

modelsummary::modelsummary(fit)
```

Equivalent Stata code:

```stata
* Tidy-like access in Stata comes from stored matrices/scalars in e()
matrix list e(b_LATE)
matrix list e(se_LATE)
matrix list e(ci_LATE)
display e(N_reps)
display e(N_misreps)
display e(share_failures)
```

## Core Stata Parity Check

The core LATE point estimates match the frozen Stata parity targets used in
the test suite:

```{r parity-check}
n_cell <- 120L

make_parity_cell <- function(g, t, ones) {
  data.frame(
    g = rep.int(g, n_cell),
    t = rep.int(t, n_cell),
    d = c(rep.int(1L, ones), rep.int(0L, n_cell - ones))
  )
}

out <- rbind(
  make_parity_cell(0L, 0L, 24L),
  make_parity_cell(0L, 1L, 48L),
  make_parity_cell(1L, 0L, 36L),
  make_parity_cell(1L, 1L, 96L)
)
out$id <- seq_len(nrow(out))
out$y <- 1 + 0.5 * out$g + 0.4 * out$t + 2 * out$d + sin(out$id / 7)
parity_fixture <- out[, c("y", "g", "t", "d")]

fit_parity <- fuzzydid(
  data = parity_fixture,
  formula = y ~ d,
  group = "g",
  time = "t",
  did = TRUE,
  tc = TRUE,
  cic = TRUE,
  breps = 5,
  seed = 1
)

stata_core <- c(
  W_DID = 2.17430877504668,
  W_TC = 2.18553887285118,
  W_CIC = 2.31982319056988
)

r_core <- setNames(fit_parity$late$estimate, fit_parity$late$estimator)[c("W_DID", "W_TC", "W_CIC")]

core_cmp <- data.frame(
  estimator = names(stata_core),
  R = unname(r_core),
  Stata = unname(stata_core),
  abs_diff = abs(unname(r_core) - unname(stata_core))
)

stopifnot(max(core_cmp$abs_diff) <= 1e-6)
knitr::kable(core_cmp, digits = 8)

# Save exact parity fixture for Stata users
if (requireNamespace("haven", quietly = TRUE)) {
  haven::write_dta(parity_fixture, "stata-parity-core-fixture.dta")
}
```

Point estimates are strict parity targets. Standard errors and confidence
intervals come from bootstrap simulation, so small cross-implementation
differences can occur even when point estimates match.

Equivalent Stata code:

```stata
* Load the same dataset exported in the parity check chunk
use "stata-parity-core-fixture.dta", clear

* Generate the group variable (already in dataset as 'g')
* The Stata command expects G to identify treatment groups

* Run fuzzydid
fuzzydid y g t d, did tc cic breps(50)

* Results are stored in e()
matrix list e(b_LATE)
matrix list e(se_LATE)
matrix list e(ci_LATE)
```

## Key Parity Points

The R package replicates Stata behavior in several important ways:

### 1. Fixed Bootstrap Seed

Both Stata and R implementations use a fixed seed (1) for bootstrap inference when standard errors are requested:

```{r seed-parity}
# These produce identical SE/CI despite different user-specified seeds
fit1 <- fuzzydid(df, y ~ d, group = "g", time = "t", did = TRUE, breps = 50, seed = 1)
fit2 <- fuzzydid(df, y ~ d, group = "g", time = "t", did = TRUE, breps = 50, seed = 999)

all.equal(fit1$late$std.error, fit2$late$std.error)

stata_seed_se <- 0.66310988000000
seed_cmp <- data.frame(
  estimator = "W_DID",
  R_seed_1 = fit1$late$std.error[fit1$late$estimator == "W_DID"],
  R_seed_999 = fit2$late$std.error[fit2$late$estimator == "W_DID"],
  Stata = stata_seed_se
)
seed_cmp$abs_diff_R1_vs_Stata <- abs(seed_cmp$R_seed_1 - seed_cmp$Stata)
seed_cmp$abs_diff_R999_vs_Stata <- abs(seed_cmp$R_seed_999 - seed_cmp$Stata)
knitr::kable(seed_cmp, digits = 8)
```

Equivalent Stata code:

```stata
use "stata-parity-sim-fixture.dta", clear
set seed 1
fuzzydid y g t d, did breps(50)
matrix se1 = e(se_LATE)

set seed 999
fuzzydid y g t d, did breps(50)
matrix se2 = e(se_LATE)

matrix list se1
matrix list se2
```

### 2. Design Restrictions

The R package enforces the same restrictions as Stata:

```{r restrictions, error = TRUE}
df_restrict <- transform(df, x = rnorm(nrow(df)))

# CIC requires no covariates
try({
  fuzzydid(df_restrict, y ~ d + x, group = "g", time = "t", cic = TRUE)
})

# LQTE requires binary G, T, and D, plus no covariates
try({
  fuzzydid(df_restrict, y ~ d + x, group = "g", time = "t", lqte = TRUE)
})
```

Equivalent Stata code:

```stata
use "stata-parity-sim-fixture.dta", clear
gen x = rnormal()

* CIC with covariates should error
capture noisily fuzzydid y g t d x, cic

* LQTE with covariates should error
capture noisily fuzzydid y g t d x, lqte
```

### 3. Partial Identification Bounds

For the Wald-TC estimator, when no stable control group exists, partial identification bounds can be computed:

```{r partial}
fit_partial <- fuzzydid(
  data = df,
  formula = y ~ d,
  group = "g",
  time = "t",
  tc = TRUE,
  partial = TRUE,
  breps = 50
)

knitr::kable(fit_partial$late)

partial_gold <- read_parity_golden("stata-parity-partial-golden.csv")
stata_partial <- stats::setNames(partial_gold$estimate, partial_gold$estimator)
r_partial <- setNames(fit_partial$late$estimate, fit_partial$late$estimator)[names(stata_partial)]
partial_cmp <- data.frame(
  estimator = names(stata_partial),
  R = unname(r_partial),
  Stata = unname(stata_partial),
  abs_diff = abs(unname(r_partial) - unname(stata_partial))
)
stopifnot(max(partial_cmp$abs_diff) <= 1e-6)
knitr::kable(partial_cmp, digits = 8)
```

Equivalent Stata code:

```stata
use "stata-parity-sim-fixture.dta", clear
fuzzydid y g t d, tc partial breps(50)
matrix list e(b_LATE)
```

### 4. Equality Tests

When multiple estimators are requested, equality tests can be performed:

```{r eqtest}
fit_eq <- fuzzydid(
  data = parity_fixture,
  formula = y ~ d,
  group = "g",
  time = "t",
  did = TRUE,
  tc = TRUE,
  cic = TRUE,
  eqtest = TRUE,
  breps = 5,
  seed = 1
)

eq_gold <- read_parity_golden("stata-parity-eqtest-golden.csv")
eq_cmp <- data.frame(
  contrast = eq_gold$contrast,
  R = fit_eq$eqtest$estimate,
  Stata = eq_gold$estimate,
  abs_diff = abs(fit_eq$eqtest$estimate - eq_gold$estimate)
)
stopifnot(max(eq_cmp$abs_diff) <= 1e-6)
knitr::kable(eq_cmp, digits = 8)
```

Equivalent Stata code:

```stata
use "stata-parity-core-fixture.dta", clear
fuzzydid y g t d, did tc cic eqtest breps(5)
matrix list e(b_LATE_eqtest)
```

# Covariate Adjustment

## Parametric Adjustment with modelx

When covariates are included, conditional expectations can be estimated via OLS, logit, or probit:

```{r covariates}
# Add covariates
set.seed(123)
df$x1 <- rnorm(nrow(df))
df$x2 <- factor(ifelse(runif(nrow(df)) > 0.5, "a", "b"))

fit_modelx <- fuzzydid(
  data = df,
  formula = y ~ d + x1 + x2,
  group = "g",
  time = "t",
  did = TRUE,
  tc = TRUE,
  modelx = c("ols", "ols"),
  nose = TRUE
)

knitr::kable(fit_modelx$late)

stata_modelx <- c(
  W_DID = 2.03222780000000,
  W_TC = 1.93062450000000
)
r_modelx <- setNames(fit_modelx$late$estimate, fit_modelx$late$estimator)[names(stata_modelx)]
modelx_cmp <- data.frame(
  estimator = names(stata_modelx),
  R = unname(r_modelx),
  Stata = unname(stata_modelx),
  abs_diff = abs(unname(r_modelx) - unname(stata_modelx))
)
knitr::kable(modelx_cmp, digits = 8)

if (requireNamespace("haven", quietly = TRUE)) {
  haven::write_dta(df[, c("y", "g", "t", "d", "x1", "x2")], "stata-parity-covariates-fixture.dta")
}
```

For binary treatments, `modelx` requires two entries:

1. Method for E(Y_{gt}|X) and E(Y_{dgt}|X)
2. Method for E(D_{gt}|X)

For multi-valued treatments, a third entry specifies the method for P(D_{gt}=d|X).

Equivalent Stata code:

```stata
use "stata-parity-covariates-fixture.dta", clear
fuzzydid y g t d, did tc continuous(x1) qualitative(x2) modelx(ols ols)
matrix list e(b_LATE)
```

## Nonparametric Adjustment with Sieves

Continuous covariates can be handled nonparametrically:

```{r sieves}
fit_sieves <- fuzzydid(
  data = df,
  formula = y ~ d + x1,
  group = "g",
  time = "t",
  did = TRUE,
  tc = TRUE,
  sieves = TRUE,
  nose = TRUE
)

knitr::kable(fit_sieves$late)

stata_sieves <- c(
  W_DID = 2.00081280000000,
  W_TC = 1.93678270000000
)
r_sieves <- setNames(fit_sieves$late$estimate, fit_sieves$late$estimator)[names(stata_sieves)]
sieves_cmp <- data.frame(
  estimator = names(stata_sieves),
  R = unname(r_sieves),
  Stata = unname(stata_sieves),
  abs_diff = abs(unname(r_sieves) - unname(stata_sieves))
)
knitr::kable(sieves_cmp, digits = 8)
```

When `sieves = TRUE`, leaving `sieveorder = NULL` uses deterministic 5-fold cross-validation (Stata-parity default). You may also pass a scalar `sieveorder` to set a fixed order for both outcome and treatment sieve bases.

Equivalent Stata code:

```stata
use "stata-parity-covariates-fixture.dta", clear
fuzzydid y g t d, did tc continuous(x1) sieves
matrix list e(b_LATE)
```

# LQTE Estimation

For quantile treatment effects at the 5th through 95th percentiles:

```{r lqte}
lqte_gold <- read_parity_golden("stata-parity-lqte-golden.csv")
fit_lqte <- fuzzydid(
  data = parity_fixture,
  formula = y ~ d,
  group = "g",
  time = "t",
  lqte = TRUE,
  nose = TRUE,
  seed = 1
)

lqte_cmp <- data.frame(
  quantile = lqte_gold$quantile,
  R = fit_lqte$lqte$estimate,
  Stata = lqte_gold$estimate,
  abs_diff = abs(fit_lqte$lqte$estimate - lqte_gold$estimate)
)
stopifnot(max(lqte_cmp$abs_diff) <= 1e-6)
knitr::kable(head(lqte_cmp, 10), digits = 8)
```

Note: LQTE requires binary G, T, and D with no covariates.

Equivalent Stata code:

```stata
use "stata-parity-core-fixture.dta", clear
fuzzydid y g t d, lqte
matrix list e(b_LQTE)
```

# Multi-Period Designs

For designs with more than two periods, use `group_forward`:

```{r multiperiod}
set.seed(7)
n_mp <- 70

make_mp_cell <- function(gb, gf, t, prob, shift = 0) {
  d <- rbinom(n_mp, 1, prob)
  y <- 1 + 0.35 * t + 1.6 * d + shift + rnorm(n_mp, sd = 0.25)
  data.frame(y = y, d = d, gb = gb, gf = gf, t = t)
}

df_mp <- rbind(
  make_mp_cell(gb = 0, gf = 0, t = 0, prob = 0.20),
  make_mp_cell(gb = 0, gf = 0, t = 1, prob = 0.30),
  make_mp_cell(gb = 0, gf = 0, t = 2, prob = 0.35),
  make_mp_cell(gb = 0, gf = 1, t = 0, prob = 0.30, shift = 0.2),
  make_mp_cell(gb = 1, gf = 1, t = 1, prob = 0.60, shift = 0.8),
  make_mp_cell(gb = 1, gf = 0, t = 2, prob = 0.75, shift = 1.1)
)

fit_mp <- fuzzydid(
  data = df_mp,
  formula = y ~ d,
  group = "gb",
  group_forward = "gf",
  time = "t",
  did = TRUE,
  tc = TRUE,
  nose = TRUE
)

knitr::kable(fit_mp$late)

if (requireNamespace("haven", quietly = TRUE)) {
  haven::write_dta(df_mp, "stata-parity-multiperiod-fixture.dta")
}
```

Equivalent Stata code:

```stata
use "stata-parity-multiperiod-fixture.dta", clear
fuzzydid y gb t d, groupf(gf) did tc
matrix list e(b_LATE)
```

# Bootstrap Diagnostics

The package tracks bootstrap failures (degenerate resamples):

```{r bootstrap-diag}
df_diag <- parity_fixture
df_diag$cl <- rep(seq_len(24L), each = 20L)
cluster_gold <- read_parity_golden("stata-parity-cluster-golden.csv")

fit_diag <- fuzzydid(
  data = df_diag,
  formula = y ~ d,
  group = "g",
  time = "t",
  did = TRUE,
  tc = TRUE,
  cic = TRUE,
  cluster = "cl",
  breps = 20,
  seed = 1
)

diag_cmp <- data.frame(
  metric = c("N.reps", "N.misreps", "Share.failures"),
  R = c(fit_diag$n_reps, fit_diag$n_misreps, fit_diag$share_failures),
  Stata = c(cluster_gold$n_reps[[1]], cluster_gold$n_misreps[[1]], cluster_gold$share_failures[[1]])
)

stopifnot(abs(diag_cmp$R[3] - diag_cmp$Stata[3]) <= 1e-12)
stopifnot(diag_cmp$R[1] == diag_cmp$Stata[1], diag_cmp$R[2] == diag_cmp$Stata[2])
knitr::kable(diag_cmp, digits = 8)

if (requireNamespace("haven", quietly = TRUE)) {
  haven::write_dta(df_diag, "stata-parity-cluster-fixture.dta")
}
```

Equivalent Stata code:

```stata
use "stata-parity-cluster-fixture.dta", clear
fuzzydid y g t d, did tc cic cluster(cl) breps(20)
display e(N_reps)
display e(N_misreps)
display e(share_failures)
```

# References

de Chaisemartin, C. and D'Haultfoeuille, X. (2018). [Fuzzy Differences-in-Differences](https://academic.oup.com/restud/article/85/2/999/4096388). *Review of Economic Studies*, 85(2): 999-1028.
